---
title: "Introduction to DSTget"
author: "Olav Grøndal"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

DSTget is a small wrapper around the Statistics Denmark statbank API. It allows for easy import of all tables from the statbank.
It has the following features.

* Variable selection and user input checking
* Circumvention of standard download limit of the statbank
* Conversion of statistical timeperiods to date
* Easy conversion to R data.frame

## Getting table meta data
First load the package
```{r}
library(DSTget)
```

To get started download the metadata of a table into a table object by using the `DSTget` function.
```{r}
table <- DSTget("BEV3C")
```
This above command download all the metadat of the table, this can be used as input for the getData function.
You can also use `summary` to get a description of the table metadata.

```{r}
summary(table)
```

## Downloading table data
For some table you can simply download them in their entirety by passing the meta data object to the getData function.

```{r}
myData <- getData(table)
```

Sometimes you want only a subset of variables or time periods, or the table might be too big to be downloaded in its
entirety. You can then select the correct variables by choosing what levels of the categorical values you want to include.

```{r}
myData <- getData(table, BEVÆGELSEV=c(18),Tid=c("2015M01","2015M02","2015M03","2015M04"))
```

You can see all the possible values for all variables in the table object.
By running `table$values`

```{r}
# Here wrapped in str() to avoid long output
str(table$values)
```

## Advanced variable selection

There are a few extra options for selecting your data.

* Using R dates
* Using fillRemaining

The first option lets you use the startDate and endDate parameters to select time periods.

```{r}
myData <- getData(table, BEVÆGELSEV=c(18), startDate = as.Date("2012-01-01"))
```
This will get all timeperiods from the startDate, I can specify a range by including both startDate and endDate, or only have an endDate to get historic data.

The other option is `fillRemaining` which lets you get all levels of all the variables you dont mention.
You might only care about 15 to 19 year olds in the highest completed education table (HFUDD10), but you want all other variables to be included
without specifically typing them in.

```{r}
table <- DSTget("HFUDD10")
myData <- getData(table, ALDER=c("15-19"), KØN = c('K'), startDate = as.Date("2015-01-01"), fillRemaining=T)
```

## Date variables
All tables from the statbank use statistical periods, like 2015M01 ie. the first month of 2015.
DSTget conveniently provides an R start and end date for each period, making it easy
to construct time series and sort and use the data in other ways.

## Downloading very large tables
By default the statbank gives you an error if you ask for tables above 100.000 rows.
DSTget will give you an error if you specify such a request. However - if you
set the `splitLarge` argument to TRUE, then DSTget will split up your request for you
and let you download the table (this might take a while).

```{r}
table <- DSTget("FOLK1A")
myData <- getData(table, ALDER = c(0:6), fillRemaining=T,startDate = as.Date("2010-01-01"), splitLarge=T)
```
